---
title: "Lecture 5: Data wrangling"
subtitle: "EC7412 Part II: Data Science for Economists"
date: 2025-04-25
---

# Introduction {background-color="#40666e"}

```{r}
#| echo: false
#| message: false
#| warning: false
# Basic setup chunk

# Load necessary packages
library(data.table)
library(dplyr)      # For comparison
library(tidyr)      # For comparison
library(haven)      # For Stata files
library(readxl)     # For Excel files
library(lubridate)  # For dates
library(stringr)    # For string operations if needed

# Set data.table print options for slides
options(datatable.print.nrows = 5)
options(datatable.print.topn = 3)

options(width = 300) #default 80
```

## The data workflow revisited

:::: {.columns}
::: {.column width="30%"}
```{mermaid}
%%| mermaid-format: png
graph TB;
    A["(1) Obtain Data"] --> B["(2) Wrangling
                              Clean, Tidy, Transform"];
    B --> C["(3) Exploration &
             Visualization"];
    C --> B;
    C --> D["(4) Analysis"];
    D --> F["(5) Communication"];
```
:::
::: {.column width="70%" .fragment}
- **Obtain** and import data
- **Wrangling=** make data usable for analysis
    - **Clean** data inconsistensies
    - **Tidy** data format
    - **Transform** data, create variables
:::
::::

## Why is wrangling important?

- Raw data is rarely ready for analysis or visualization
- Errors, inconsistencies, and structural issues can lead to:
    - Incorrect results
    - Wasted time debugging later
    - Difficulty merging or comparing datasets
- Good wrangling practices ensure data quality and make analysis smoother
- *Garbage In, Garbage Out*

## Our r tools: `data.table` and `dplyr`

- Two dominant paradigms for data manipulation in R
- Both are powerful and widely used, but with different syntax
- We will focus on `data.table`, because it is much faster
- Show `dplyr` equivalents for comparison

## `data.table`: Core concept

- An **enhancement** of base R's `data.frame`
- Designed for **speed** and **memory efficiency**
- Concise, general syntax:

  ```r
  DT[i, j, by]
  ```

  - `DT`: your data.table object
  - `i`: Row filtering (SQL `WHERE`)
  - `j`: Column operations (SQL `SELECT`, `UPDATE`)
  - `by`: Grouping operation (SQL `GROUP BY`)

## `data.table`: Key features

- **Fast:** Operations implemented in C
- **Memory efficient:** Modifies *by reference* using `:=`^[But this means that when you do not want to edit by reference you have to be explicit and use `copy()`.]
- **Indexing:** Automatic indexing (with keys) speeds up subsets and joins
- **Concise syntax:** Combines operations efficiently

## `dplyr`: Core concept

- Part of the **Tidyverse** collection of packages
- Focuses on **readability** and **consistency**
- Uses distinct "verbs" for common operations

## `dplyr`: Key verbs

- `filter()`: Subset rows
- `select()`: Select columns
- `mutate()`: Create/modify columns
- `summarise()` / `summarize()`: Aggregate data
- `group_by()`: Perform operations group-wise
- `arrange()`: Reorder rows


## Example syntax

:::: {.columns}
::: {.column width="50%"}
`data.table`
```r
data[condition,
     .(new_var = mean(old_var)),
     by = "group_var"]
```
:::
::: {.column width="50%"}
`dplyr`
```r
data |>
  filter(condition) |>
  group_by(group_var) |>
  summarise(new_var = mean(old_var))
```
:::
::::

::: {.aside}
If you really like the `dplyr` syntax but want to take advantage of `data.table` speeds, there is the [dtplyr](https://dtplyr.tidyverse.org/) package which uses data.table as a backend for dplyr-syntax operations.
:::

## Key differences

**`data.table`**

- **Pros:** Speed, memory efficiency, concise syntax, powerful indexing/joins. Ideal for large data.
- **Cons:** Steeper learning curve, syntax less "English-like", modify-by-reference needs care.

**`dplyr`**

- **Pros:** Highly readable verbs, Tidyverse integration (ggplot2, etc.).
- **Cons:** Slower, uses more memory, especially on large data or complex groups.


# Reading data {background-color="#40666e"}

## `data.table::fread()` for text files

- `fread()` is fast and flexible for delimited text (CSV, TSV)
- Automatically solves many possible data issues
- Let's try to read some AI-generated data:

::: {.fragment}
```{r}
#| error: true
#| warning: true
#| output-location: default
dogs = fread("dog_inventory.csv")
```

::: {.small-output}
```{r}
#| output-location: default
dogs
```
:::
:::

## Fixing issues: character encoding

- Computers need to encode characters to store them
- Unfortunately there are tons of different encodings
- A-Z, 0-9 is usually the same, but code for e.g. å can vary
- Can create a real mess...

::: {.small-output .fragment}
```{r}
#| warning: false
#| output-location: default
fread("dog_inventory.csv", encoding = "UTF-8", skip = 174, nrows = 1)
fread("dog_inventory.csv", encoding = "Latin-1", skip = 174, nrows = 1)
```
:::

## `fread()`: fixing names

```{r}
#| output-location: default
dogs = fread("dog_inventory.csv", encoding = "Latin-1")
names(dogs)
```

::: {.fragment}
To refer to a name with a space we need to enclose it with backticks (\`\`)

```{r}
#| output-location: default
dogs$`new owner`[5]
```
:::

::: {.fragment}
`check.names=TRUE` fixes column names automatically

```{r}
#| output-location: default
names(fread("dog_inventory.csv", check.names = TRUE))[10]
```
:::

## `fread()`: coding missing data

```{r}
#| output-location: default
unique(dogs$insured)
```

`na.strings` argument takes vector of strings that should be coded as `NA`

```{r}
#| output-location: default
dogs = fread("dog_inventory.csv",
             encoding = "Latin-1",
             check.names = TRUE,
             colClasses = c(dog_id = "character"),
             na.strings = c("", "-99"))
```

## `fread()`: selecting columns

Use `select` to read a subset of columns


::: {.small-output}
```{r}
#| output-location: default
fread("dog_inventory.csv",
      select = c("dog_id", "name", "breed"))
```
:::

## `fread()`: specifying types

`fread()` tries to guess the data types

- override auto-detection for problematic columns with `colClasses`:

::: {.small-output}
```{r}
#| output-location: default
fread("dog_inventory.csv",
      select = c("dog_id", "name", "breed"),
      colClasses = c(dog_id = "character"))
```
:::

::: {.aside}
You can also use `select` to both chose columns and specify types by supplying a named vector, see `?fread`.
:::

## Reading directly from urls

`fread()` handles reading text files directly from the web

::: {.small-output}
```{r}
#| output-location: default
fread("https://github.com/adamaltmejd/datascience-course/raw/refs/heads/main/lectures/lecture_5/dog_inventory.csv")
```
:::

## Importing Stata `.dta` files with `haven`

- Use `haven::read_dta()`. Preserves Stata labels.
- Convert to `data.table` with `as.data.table()`,
- Set value labels with `as_factors()`

```r
library(haven)

stata_dt = read_dta("path/to/stata_file.dta") |>
  as.data.table()

stata_dt[, factor_variable := as_factor(labelled_variable)]
```

::: {.aside}
Similarly, we would use `haven::write_dta()` to save a data frame as a Stata dta file.
:::

## Reading excel files with `readxl`

```r
library(readxl)
excel_dt = read_excel("path/to/spreadsheet.xlsx")
```

`readxl` supports selecting sheets:

```r
excel_sheets("path/to/spreadsheet.xlsx")
read_excel("spreadsheet.xlsx", sheet = "Sheet2")
```

and even specific ranges:

```r
read_excel("spreadsheet.xlsx", range = "A1:D10", skip = 2)
```

# Subsetting / Filtering (i) {background-color="#40666e"}

## Subsetting / Filtering

`DT[condition, ...]` and `dplyr::filter(condition)` subsets rows based on `condition`.

Selecting dogs aged less than 0:

::: {.panel-tabset .small-output}
#### `data.table`

```{r}
#| output-location: default
dogs[age < 0]
```

#### `dplyr`

```{r}
#| output-location: default
filter(dogs, age < 0)
```
:::

## Subsetting / Filtering (cont.)

Selecting insured unsold dogs:

::: {.panel-tabset .small-output}
#### `data.table`

```{r}
#| output-location: default
dogs[insured == 1 & is.na(sale_date)]
```

#### `dplyr`

```{r}
#| output-location: default
filter(dogs, insured == 1 & is.na(sale_date))
```
:::

## Selecting using row numbers

Select first and last row (note `.N` and `row_number()`):

::: {.panel-tabset .small-output}
#### `data.table`

```{r}
#| output-location: default
dogs[c(1,.N)]
```

#### `dplyr`

```{r}
#| output-location: default
filter(dogs, row_number() %in% c(1, n()))
```
:::

Exclude final 10 rows:

::: {.panel-tabset .small-output}
#### `data.table`

```{r}
#| output-location: default
dogs[1:(.N-11)]
```

#### `dplyr`

```{r}
#| output-location: default
filter(dogs, row_number() < n()-10)
```
:::

# Working with columns (j) {background-color="#40666e"}

## Working with columns (j)

- `data.table` syntax `DT[i, j, by]`: operates on columns with `j`.
- Use `.()` or `list()` to output a new data.table.
- `dplyr::select()` works similarly

::: {.panel-tabset .small-output}
#### `data.table`

```{r}
#| output-location: default
dogs[, .(name)]
```

#### `dplyr`

```{r}
#| output-location: default
select(dogs, name)
```
:::

## Excluding columns

::: {.panel-tabset .small-output}
#### `data.table`

```{r}
#| output-location: default
dogs[, !c("sale_date", "new.owner")]
```

#### `dplyr`

```{r}
#| output-location: default
select(dogs, !c(sale_date, new.owner))
```
:::

## `data.table` assignment by reference (`:=`)

- Normally to edit a column, R copies the whole dataset in memory
- But `:=` modifies `DT` **in place**

::: {.panel-tabset .small-output}
#### `data.table`

```{r}
#| output-location: default
dogs[, price := price * 10]
```

#### `dplyr`

```{r}
#| output-location: default
dogs = mutate(dogs, price = price / 10)
```
:::

::: {.aside}
`copy(DT)` first if you *want* to work with a copy of `DT`
:::


## Removing a column by reference

To delete a `data.table` column, assign it to `NULL`.

```{r}
#| output-location: default
dogs[, new.owner := NULL]
```


## Conditional modification with `:=`

- Combine row selection (`i`) with assignment (`j`) for conditional modification in `data.table`
- Here is where data.table is starting to shine

::: {.panel-tabset .small-output}
#### `data.table`

```{r}
#| output-location: default
#| error: true
dogs[age < 0, age := NA]
```

#### `dplyr`

```{r}
#| output-location: default
dogs = mutate(dogs, age = if_else(age < 0, NA, age))
```
:::

::: {.aside}
Changing the content of a column in dplyr is always done on a copy, so we need to reassign it for the change to stay.
:::


# Grouping and summarizing (by) {background-color="#40666e"}

## Grouping and summarizing

- `DT[i, j, by = ...]` performs operation `j` separetely by group
- Corresponds to `dplyr::group_by() |> ...`

::: {.panel-tabset .small-output}
#### `data.table`

```{r}
#| output-location: default
dogs[, mean(price), by = .(insured)]
```

#### `dplyr`

```{r}
#| output-location: default
dogs |>
  group_by(insured) |>
  summarize(mean(price))
```
:::


## Grouping and summarizing (cont.)

Can group by multiple variables, and by conditions:

::: {.panel-tabset .small-output}
#### `data.table`

```{r}
#| output-location: default
dogs[, mean(price), by = .(insured, color == "White")]
```

#### `dplyr`

```{r}
#| output-location: default
dogs |>
  group_by(insured, color == "White") |>
  summarize(mean(price))
```
:::

## `.SD`

Inside a `data.table`, `.SD` refers to the **S**ubset of **D**ata for each group.

We can use this to run more complex operations, like

```r
dogs[, lapply(.SD, func), by = ...]
```

applying a custom function `func` to each `.SD` group.

::: {.aside}
`.SD` is also a data.table
:::

## `.SDcols`

Specifies the columns included in `.SD`. Can be used to select columns in smart ways.

The `dplyr` solution is simpler though (if you know the verb).

::: {.panel-tabset .small-output}
#### `data.table`
```{r}
#| output-location: default
dogs[, .SD, .SDcols = patterns("_date")]
```
#### `dplyr`
```{r}
#| output-location: default
select(dogs, ends_with("_date"))
```
:::

::: {.aside}
`patterns()` filters by *regular expression*
:::

## Save grouped summaries to column

Saving the average price by age to a new column, repeating the summary for each row in group:

::: {.panel-tabset .small-output}
#### `data.table`

```{r}
#| output-location: default
dogs[,
     avg_price_by_age := mean(price),
     by = .(age)]
```

#### `dplyr`

```{r}
#| output-location: default
dogs_new = dogs |>
  group_by(age) |>
  mutate(avg_price_by_age = mean(price))
```
:::

&nbsp;

::: {.small-output}
```{r}
dogs[order(age),
     .(age, avg_price_by_age)]
```
:::

```{r}
#| echo: false
#| output-location: default
dogs[, avg_price_by_age := NULL]
```

## Grouping and summarizing (by)

Find cheapest dog by breed, ordered by how frequent the breed is

::: {.panel-tabset}

#### `data.table`

```{r}
#| output-location: default
dogs[, .(.N, min(price)), by = "breed"] |>
  _[order(-N)]
```

#### `dplyr`

```{r}
#| output-location: default
dogs |>
  group_by(breed) |>
  summarize(N=n(), min(price)) |>
  ungroup() |>
  arrange(-N)
```

:::

# Working with strings and dates {background-color="#40666e"}

## Working with strings and dates

- Text data often requires cleaning and manipulation.
- Common tasks: detecting/replacing patterns, changing case, removing white space
- LLMs can help with some things, but are **not deterministic**


## Working with strings and dates

Convert case, remove whitespace

```{r}
tolower("Foo Bar")
```

```{r}
toupper("Foo Bar")
```

```{r}
trimws("  Foo Bar ")
```

Substring by position

```{r}
substr("Foo Bar", 4, 8)
```

Count characters^[Not to be mistaken by `length()` which returns the length of the vector, i.e. `1` for a singleton.]

```{r}
nchar("Foo Bar")
```

## Regular expressions

Regular expressions = special characters that match strings

```{r}
s = c("abc", "abcd", "abcde")
```

```{r}
#| output-location: column-fragment
grepl("abc", s)
```

```{r}
#| output-location: column-fragment
grepl("de", s)
```

```{r}
#| output-location: column-fragment
grepl("d$", s)
```

```{r}
#| output-location: column-fragment
grepl("^a", s)
```

```{r}
#| output-location: column-fragment
grepl("^[a-z]{3}$", s)
```

## Regular expressions: common tokens

- `.` matches any single character.
- `*` matches the previous item zero or more times
- `\\s` matches whitespace
- `[...]` specify possible characters

::: {.small-output}
```{r}
#| output-location: default
dogs[grepl("[Bb]rindle", color), .(name, breed, color)]
```
:::

::: {.aside}
Check `?regex` for an intro to regular expressions in R, <https://regexone.com> for a site with simple exercises to help you learn. Chatbots are pretty good at writing regexps, but do test their output!
:::

## Detecting patterns: `grepl`

- `grepl("pattern", text)` checks if `text` contains the `pattern`.
- Returns `TRUE`/`FALSE`

```{r}
dogs[grepl("Terrier", breed) &
     age <= 2, .(name, breed, price)]
```

```{r}
dogs[grepl("and", color), .N, color] |>
  _[order(-N)][1:3]
```

## Replacing patterns: `gsub`

- `gsub("pattern", "replacement", text)` replaces *all* occurrences of `pattern` with `replacement` in `text`.

Replace "and" with comma in color
```{r}
dogs[, color := gsub("and ", ",", color)]
```

Remove anything after space in dog names

```{r}
dogs[grepl(" ", name), name]
```
```{r}
dogs[grepl(" ", name),
     gsub(" .*$", "", name)]
```

::: {.aside}
Use `sub()` instead to replace only the *first* occurrence.
:::

## `stringr`

The `stringr` package provides lots of convenience functions for more powerful string operations.

- `str_replace_all()` allows for multiple patterns
- `str_to_lower()`, `str_to_upper()`, `str_trim()` are more robust
- `str_sub` permits negative indices

## Date types

R has built-in **date data types**, enabling e.g., arithmetic operations on dates

```{r}
#| error: true
"2024-02-28" + 3
```

```{r}
as.Date("2024-02-28") + 3
```
```{r}
as.Date("2024-02-28") - as.Date("2022-02-28")
```

## Date types

Let's look at the two date-columns in our data

```{r}
dogs[1:3, .(intake_date, sale_date)]
```

One is loaded as type `IDate` other as `char`. Only `IDate` is a date type:

```{r}
#| error: true
#| output-location: default
dogs[, mean(sale_date - intake_date, na.rm = TRUE)]
```

## Coercing dates

We can use `as.Date()` to coerce

```{r}
#| error: true
dogs[, sale_date := as.IDate(sale_date)]
```

::: {.fragment}
... R cannot read the format, we need to specify it manually:
```{r}
#| output-location: default
dogs[, sale_date := as.IDate(sale_date, format = "%m/%d-%Y")]
```
:::

::: {.fragment}
Then we can calculate average time in the shelter:

```{r}
#| output-location: default
dogs[, mean(sale_date - intake_date, na.rm = TRUE)]
```
:::


::: {.aside}
We need to use `as.IDate()` and not `as.Date()` because data.table defaults to the IDate format since it uses less memory. We could also have changed `intake_date` to `Date` and used `as.Date(sale_date)`.
:::

## `lubridate` package

For more advanced date operations.

```{r}
#| output-location: default
library(lubridate)
dogs[, .N, lubridate::quarter(intake_date)]
```

```{r}
date("2024-12-13") - weeks(3)
```

```{r}
hms("11:15:00") - hms("11:00:00")
```

And much more!

::: {.aside}
Lubridate cheat sheet: <https://rawgit.com/rstudio/cheatsheets/main/lubridate.pdf>
:::

# Merging and reshaping {background-color="#40666e"}

## The data

Let's load some more LLM-generated data on dogs:

::: {.small-output}
```{r}
#| output-location: default
dog_breeds = fread("dog_breeds.csv")
dog_breeds
```

```{r}
#| output-location: default
dog_sales = fread("dog_sales.csv")
dog_sales
```
:::

## Merging

- Combines two tables based on matching **keys** in shared column(s)
- Four types:
    - **Inner Join:** Keep only rows with matching keys in *both* tables
    - **Left outer join:** Keep left table, add matches from the right
    - **Right outer join:** Keep right table, add matches from the left
    - **Full outer join:** Keep all rows from *both* tables

## Merging syntax overview (SQL)

`data.table` has two alternatives

- `merge(x, y, by = "key")`
- `Y[X, on = "key"]` more efficient, allows calculations during join

`dplyr` is more verbose

- Inner join: `inner_join(x, y, by = "key")`
- Outer joins:
  - `left_join(x, y, by = "key")`
  - `right_join(x, y, by = "key")`
  - `full_join(x, y, by = "key")`


## Inner join

- Keep rows that have matching keys in *both* tables
- Show inventory data for dogs that appear in breed info table

::: {.panel-tabset .tiny-output}

#### `data.table`

```{r}
#| output-location: default
# nomatch = NA is default
dog_breeds[dogs, on = c("breed"), nomatch = NULL]
```

#### `merge.data.table`

```{r}
#| output-location: default
# all = FALSE is the default
merge(dogs, dog_breeds, by = "breed", all = FALSE, sort = FALSE)
```

#### `dplyr::inner_join`

```{r}
#| output-location: default
dogs |>
  inner_join(dog_breeds, by = "breed")
```
:::

::: {.aside}
We need to add `sort = FALSE` to `merge()` or the output will be sorted by the key column.
:::

## Left outer join

- Keep all rows from the main table and add data from the new table
- Rows in the main table with no match get `NA`
- Add information about the breeds in our inventory, keeping track of the whole inventory even if there is no breed info

::: {.panel-tabset .tiny-output}
#### `data.table`
```{r}
#| output-location: default
# nomatch = NA is default
dog_breeds[dogs, on = c("breed"), nomatch = NA]
```
#### `merge.data.table`
```{r}
#| output-location: default
# all.x = all = FALSE is default
merge(dogs, dog_breeds, by = "breed", all.x = TRUE, sort = FALSE)
```
#### `dplyr::left_join`
```{r}
#| output-location: default
dogs |>
  left_join(dog_breeds, by = "breed")
```
:::

## Right outer join

- Same as left outer join but we keep all *new* data
- Keep inventory data only for dog breeds that we have info about, also include dog info for breeds not in our kennel

::: {.panel-tabset .tiny-output}
#### `data.table`
```{r}
#| output-location: default
# nomatch = NA is default
dogs[dog_breeds, on = c("breed"), nomatch = NA]
```
#### `merge.data.table`
```{r}
#| output-location: default
# all.x = all = FALSE is default
merge(dogs, dog_breeds, by = "breed", all.y = TRUE, sort = FALSE)
```
#### `dplyr::left_join`
```{r}
#| output-location: default
dogs |>
  right_join(dog_breeds, by = "breed")
```
:::

## Full outer join

- Keep all rows from *both* tables
- Fill with `NA` when there is no match
- Keep whole inventory but add breed info even for dogs not in inventory

::: {.panel-tabset .tiny-output}
#### `merge.data.table`
```{r}
#| output-location: default
merge(dogs, dog_breeds,
      by = "breed", all = TRUE, sort = FALSE)
```
#### `dplyr::full_join`
```{r}
#| output-location: default
dogs |>
  full_join(dog_breeds, by = "breed")
```
:::

:::aside
`data.table`'s `X[Y]` syntax doesn't have a direct equivalent for `full_join`. Use `merge()`.
:::

## Data table specials: merging and manipulating

- The `X[Y]` operation allows concurrent manipulation of vectors
- Uses only needed columns, meaning we can add columns by reference
- Use `i.` syntax to refer to a column in `i`

::: {.small-output}
```{r}
#| output-location: default
dogs[dog_breeds, on = c("breed"), avg_years_left := i.avg_life_exp - age]
dogs
```
:::

## Transforming data: wide↔️long

- Often needed to make data *tidy* and easy to work with
- Remember:
    1. Each variable forms a column
    2. Each observation forms a row
    3. Each cell holds a single value

## Wide to Long

- Combines columns and increases rows
- Let's create a table of breed stats

::: {.panel-tabset}

#### `data.table::melt`
```{r}
#| output-location: default
melt(dog_breeds,
     id.vars = c("breed"),
     measure.vars = c("height_cm_min", "height_cm_max"))
```

#### `tidyr::pivot_longer`
```{r}
#| output-location: default
dog_breeds |>
  pivot_longer(cols = c("height_cm_min", "height_cm_max")) |>
  select(breed, name, value)
```

::: {.aside}
The data.table [vignette on reshaping](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-reshape.html).
:::

:::

## Reshaping: Long to Wide

- Spread values from a column across multiple new columns.
- Create a table of average sale price per breed at each shelter

::: {.panel-tabset .small-output}
#### `data.table::dcast`
```{r}
#| output-location: default
dog_sales[, .(avg_price = mean(price_usd, na.rm = TRUE)), by = .(breed, shelter_name)] |>
  dcast(breed ~ shelter_name,
        value.var = "avg_price")
```
#### `tidyr::pivot_wider`
```{r}
#| output-location: default
dog_sales |>
  group_by(breed, shelter_name) |>
  summarise(avg_price = mean(price_usd, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = shelter_name,
              values_from = avg_price)
```
:::

## Final remarks {.title-slide background-color="#40666e"}

## Data table vs Tidyverse

`tidyverse`/`dplyr` and `data.table` are all great.

While `tidyverse` is easy to get started with and understand, `data.table` is much more powerful.

Use whichever you prefer!

::: {.aside}
And if you really like `dplyr` syntax but want `data.table` speeds, check out the [`dtplyr`](https://dtplyr.tidyverse.org/) package which provides a data.table backend for lots of dplyr commands.
:::

## Data table magic

I'm a bit of a `data.table` nerd though, so here is a cool examples of advanced operation.

Let's calculate the average price during two months around each sale:

```{r}
#| output-location: default
#| code-line-numbers: "3-6|7-9|1-12"
dogs[
  dog_sales[
    dogs[!is.na(sale_date),
         .(dog_id, breed,
           date1 = as.Date(sale_date) %m-% months(2),
           date2 = as.Date(sale_date) %m+% months(2))],
    on = c("breed", "date >= date1", "date < date2"),
    by = .EACHI,
    .(dog_id, mean_price = mean(price_usd, na.rm = TRUE))
  ],
  on = "dog_id", avg_price_2months := i.mean_price
]
```

## Data table magic (cont.)

Let's verify it worked

```{r}
#| output-location: default
dogs[dog_id == 9367, .(breed, sale_date, avg_price_2months)]
```

```{r}
#| output-location: default
dog_sales[breed == "Border Collie" &
          date %between% list(as.Date("2024-01-12") - months(2),
                              as.Date("2024-01-12") + months(2)),
          mean(price_usd)]
```

## Data table magic (cont.)

- `data.table` also supports rolling joins, where we find the *closest* match
- `roll` can go forward, backward or to nearest match
- Let's find the sale price closest to `sale_date`

```{r}
#| output-location: default
dogs[
  dog_sales[dogs, on = c("breed", "date == sale_date"), roll = "nearest", .(dog_id, price_usd)],
  on = "dog_id",
  nearest_price := i.price_usd
]
dogs[dog_id == "0013", .(breed, sale_date, nearest_price)]
dog_sales[breed == "Labrador Retriever" &
          date %between% list(as.Date("2023-03-01"), as.Date("2023-04-01"))]
```

## Saving data

You should split your data processing and analysis code.

`src/process_dogs.R` takes the raw dogs data and outputs a file `data/dogs_processed.csv`. Use `data.table::fwrite()` to save the processed data.

Why not `write_rds` or `save`? Like `.dta` these are binary formats, cannot be tracked by version control for example.

`src/analyze_dogs.R` starts by `fread(data/dogs_processed.csv)` and then runs analysis on this processed data.

## Combining operations

Both `data.table` and `dplyr` allow chaining multiple steps using **pipes**.

Great for readability!

`data.table` `DT[...] |> _[...]`, where  `_` represents the data.table from the previous step

## Next lecture: Servers, VMs, APIs, & LLMs {.title-slide background-color="#40666e"}